The Anatomy of a Temporal Summarization System

### 2015-02-28 conclusion.tex
ck rev

In the future we would like to explore the applicability of this model and features
%KM3 - I think not just for more general domains (as it says something wrong about this one) but to a wider class of events.
to a wider class of events. Additionally, we want to further analyze the failure 
of the temporal relevance features; intuitively we believe time to be important
for the update summarization task.

>>>>

In the future we would like to explore the application of the AP+Salience
 model and features
%KM3 - I think not just for more general domains (as it says something wrong about this one) but to a wider class of events.
to a wider class of events.

We would also like to further analyze the failure 
of the temporal relevance features; intuitively we believe time to be important
for the update summarization task.

### 2015-02-28 introduction.tex
ck rev

Multi-document summarization has the potential to assist the crisis 
informatics community. 
Automatic summarization could deliver
relevant and salient information at regular intervals, 
even when human volunteers are unable to. 
Perhaps more importantly it could help filter out
unnecessary and irrelevant detail when the volume of incoming information
is large. 
While methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively
%KM - I think we should consider adding other event summarizers. 
(e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}), 
these experiments were not developed to handle streaming data from a
 heterogeneous environment at web scale. 
These methods also rely heavily on redundancy which is suboptimal for time
sensitive domains where there is a high cost in delaying information.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
We demonstrate that combining salience with clustering produces more relevant summaries
compared to baselines using clustering alone.  
We suggest that this
because our system is better able to adapt to
dynamic changes in input volume which adversely 
effect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience does not just represent importance within a set of
documents; it also represents both how typical a sentence is of the  
event
type (i.e., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
expected similarity of their ideal summaries.
%While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
%expected similarity of their ideal summaries.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

>>>>

Multi-document summarization has the potential to assist the crisis 
informatics community. 
Automatic summarization could deliver
relevant and salient information at regular intervals, 
even when human volunteers are unable to. 
Perhaps more importantly it could help filter out
unnecessary and irrelevant detail when the volume of incoming information
is large. 
While methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively
%KM - I think we should consider adding other event summarizers. 
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}, 
these experiments were not developed to handle streaming data from a
 heterogeneous environment at web scale. 
These methods also rely heavily on redundancy which is suboptimal for time
sensitive domains where there is a high cost in delaying information.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
We demonstrate that combining salience with clustering produces more relevant summaries
compared to baselines using clustering alone.  
Our experiments suggest that this
because our system is better able to adapt to
dynamic changes in input volume which adversely 
effect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience does not just represent importance within a set of
documents; it also represents both how typical a sentence is of the  
event
type (e.g., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we evaluate these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
expected similarity of their ideal summaries.
%While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
%expected similarity of their ideal summaries.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

### 2015-02-28 methods.tex
ck rev

In order to measure the usefulness of a system's updates, we consider
the degree to which the system output reflects the different
aspects of an event.  Events are often composed of a variety of sub-events.  
For example, the Hurricane Sandy
event includes sub-events related to the storm making landfall,
the ensuing flooding, the many transportation issues, amongst many
others.  An ideal system would update the user about each of these
sub-events as they occur, with low latency.  
%In order to use
%language consistent with previous literature, 
We refer to these
sub-events as the \emph{nuggets} associated with an event.  Nuggets are
defined as a fine-grained atomic sub-event associated with an event.  
%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
We present several example nuggets associated with the Hurricane
Sandy event 
in Figure \ref{fig:nuggets}.  
We describe how these 
nuggets are found in Section \ref{sec:data}.

%\input{nugget_examples.tex}

Throughout our treatment of our algorithm, the \emph{salience} 
of an update captures the degree to which it reflects an 
event's nuggets.  Assuming that we have a text representation 
of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
defined as,
\begin{align}
\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
\operatorname{sim}(u, n) \label{eq:salience}
\end{align}
where $\operatorname{sim}(\cdot)$ is the semantic similarity such as
the cosine similarity of latent vectors associated with the update and 
nugget text \cite{guo2012simple}.

@@

Our system architecture follows a simple pipeline design where each
stage provides an additional level of processing or filtering of the input
sentences.
We begin with an empty update summary $U$.
At each hour we receive a new batch of sentences $b_t$ from the stream
and perform the following actions:
\begin{enumerate}
  \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
  \item select a set of exemplar sentences in $b_t$ by combining 
      clustering with 
      salience predictions (Section~\ref{sec:exsel}),
  \item add the most novel and salient exemplars 
      to $U$ (Section~\ref{sec:upsel}).
\end{enumerate}
The resultant list of updates $U$ is our summary of the event.

@@

The second model is estimated from text specific to our event types.  
For each event type we create a corpus of related documents using pages
and subcategories listed under a related Wikipedia category.
For example, the language model for event type `earthquake' is estimated 
from Wikipedia pages under the category \emph{Category:Earthquakes}.

%\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
%expected summary is similar to some previously seen target or some side information.}

These models are intended to detect sentences similar to those appearing in 
summaries of other events in the same category 
(e.g. most earthquake summaries are likely to include higher probability for 
ngrams including the token `magnitude'). While we focus our system on the 
language of news and disaster, we emphasize that the use of language modeling 
can be an effective feature for multi-document summarization for other 
domains that have related text corpora.

@@

At run-time
our summarization system
does not have access to event nuggets to compute sentence salience. 
However, we can use the nuggets
in our training data to train a model to predict salience for sentences
in new events.
For each event in our training data, we sample a set of sentences and  each 
sentence's salience is computed according to Equation \ref{eq:salience}.
This results in a training set of sentences and the target values 
(their salience) to predict.

Now that we have a feature representation for the sentences, we can use
any off the shelf regression technique to predict salience.
We opt to use a Gaussian process (GP) regression model
\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
kernel.  
Our features fall naturally into five groups and we use a separate RBF kernel
for each, using the sum of each feature group RBF kernel as the final input
to the GP model.
%GP regressors are a
%class of data-driven, non-parametric model generalizing the multi-variate
%Gaussian to the infinite dimensional setting.

@@

Once we have predicted the salience for a batch of sentences, we must
now select a set of update candidates, i.e. sentences that are both salient
and representative of the current batch.
To accomplish this, we combine the output of our 
salience prediction model with the affinity
propagation algorithm 
%to identify a set of exemplar sentences 
%for each input batch. 
Affinity propagation (AP) is a clustering algorithm
that identifies a subset of datapoints as exemplars and forms clusters
by assigning the remaining points to one of the exemplars 
\cite{frey2007clustering}. AP attempts to 
maximize the net similarity objective 
\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
+ \sum_{i : i = e_i}^n \operatorname{salience}(e_i)  \]
where $e_i$ is the exemplar of the $i$-th data point, and functions
$\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
similarity
of data points and the apriori preference of a data point to be an exemplar
respectively. 
AP differs from other $k$-centers algorithms in that it simultaneously 
considers all data points as exemplars, making it less prone to finding
local optima as a result of poor initialization. Furthermore, the 
second term in $\mathcal{S}$ incorporates the individual importance of 
data points as candidate exemplars; most other clustering algorithms only make
use of the first term, i.e. the pairwise similarities between data points.

@@

 Next,
to prevent adding updates that are redundant, we filter out exemplars
that are too similar to previous updates.
The exemplars are examined
sequencially in order of decreasing salience and  a similarity threshold 
is applied, where the exemplar is ignored if its
maximum semantic similarity to any previous updates in the summary is
greater than $\lambda_{sim}$.

>>>>

In order to measure the usefulness of a system's updates, we consider
the degree to which the system output reflects the different
aspects of an event.  Events are often composed of a variety of sub-events.  
For example, the Hurricane Sandy
event includes sub-events related to the storm making landfall,
the ensuing flooding, the many transportation issues, amongst many
others.  An ideal system would update the user about each of these
sub-events as they occur, with low latency.  
%In order to use
%language consistent with previous literature, 
We refer to these
sub-events as the \emph{nuggets} associated with an event.  A nugget is
defined as a fine-grained atomic sub-event associated with an event.  
%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
We present 2 example nuggets associated with the Hurricane
Sandy event 
in Figure \ref{fig:nuggets}. Note that each event has anywhere from 50 to
several hunder nuggets in total.
We describe how these 
nuggets are found in Section \ref{sec:data}.

\input{nugget_examples.tex}

Throughout our treatment of our algorithm, the \emph{salience} 
of an update captures the degree to which it reflects an 
event's nuggets.  Assuming that we have a text representation 
for each of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
defined as,
\begin{align}
\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
\operatorname{sim}(u, n) \label{eq:salience}
\end{align}
where $\operatorname{sim}(\cdot)$ is the semantic similarity such as
the cosine similarity of latent vectors associated with the update and 
nugget text \cite{guo2012simple}.

@@

Our system architecture follows a simple pipeline design where each
stage provides an additional level of processing or filtering of the input
sentences.
We begin with an empty update summary $U$.
At each hour we receive a new batch of sentences $b_t$ from the stream
of event relevant documents
and perform the following actions:
\begin{enumerate}[nosep]
  \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
  \item select a set of exemplar sentences in $b_t$ by combining 
      clustering with 
      salience predictions (Section~\ref{sec:exsel}),
  \item add the most novel and salient exemplars 
      to $U$ (Section~\ref{sec:upsel}).
\end{enumerate}
The resultant list of updates $U$ is our summary of the event.

@@

The second model is estimated from text specific to our event types.  
For each event type we create a corpus of related documents using pages
and subcategories listed under a related Wikipedia category.
For example, the language model for event type `earthquake' is estimated 
from Wikipedia pages under the category \emph{Category:Earthquakes}.  
%\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
%expected summary is similar to some previously seen target or some side information.}
These models are intended to detect sentences similar to those appearing in 
summaries of other events in the same category 
(e.g. most earthquake summaries are likely to include higher probability for 
ngrams including the token `magnitude'). While we focus our system on the 
language of news and disaster, we emphasize that the use of language modeling 
can be an effective feature for multi-document summarization for other 
domains that have related text corpora.

@@

At run-time
our summarization system
does not have access to event nuggets to compute sentence salience. 
However, we can use the nuggets
in our training data to model salience as a function of our features.
Such a model generalizes to new events.
For each event in our training data, we sample a set of sentences and  each 
sentence's salience is computed according to Equation \ref{eq:salience}.
This results in a training set of sentences and the target values 
(their salience) to predict.

Given a feature representation for the sentences, we can use
any regression technique to predict salience.
We opt to use a Gaussian process (GP) regression model
\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
kernel.  
Our features fall naturally into five groups and we use a separate RBF kernel
for each, using the sum of each feature group RBF kernel as the final input
to the GP model.
%GP regressors are a
%class of data-driven, non-parametric model generalizing the multi-variate
%Gaussian to the infinite dimensional setting.

@@

Once we have predicted the salience for a batch of sentences, we must
now select a set of update candidates, i.e. sentences that are both salient
and representative of the current batch.
To accomplish this, we combine the output of our 
salience prediction model with the affinity
propagation algorithm.
%to identify a set of exemplar sentences 
%for each input batch. 
Affinity propagation (AP) is a clustering algorithm
that identifies a subset of datapoints as exemplars and forms clusters
by assigning the remaining points to one of the exemplars 
\cite{frey2007clustering}. AP attempts to 
maximize the net similarity objective 
\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
+ \sum_{i : i = e_i}^n \operatorname{salience}(e_i)  \]
where $e_i$ is the exemplar of the $i$-th data point, and functions
$\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
similarity
of data points and our predicted apriori preference of a data point to be an exemplar
respectively. 
AP differs from other $k$-centers algorithms in that it simultaneously 
considers all data points as exemplars, making it less prone to finding
local optima as a result of poor initialization. Furthermore, the 
second term in $\mathcal{S}$ incorporates the individual importance of 
data points as candidate exemplars; most other clustering algorithms only make
use of the first term, i.e. the pairwise similarities between data points.

@@

 Next,
to prevent adding updates that are redundant with previous output updates, 
we filter out exemplars
that are too similar to previous updates.
The exemplars are examined
sequencially in order of decreasing salience and  a similarity threshold 
is applied, where the exemplar is ignored if its
maximum semantic similarity to any previous updates in the summary is
greater than $\lambda_{sim}$.

### 2015-02-28 relatedwork.tex
ck rev

First, centrality-focused approaches (including graph,
cluster, and centroid methods) are very natural for retrospective analysis in
the sense that they let the data "speak for itself." 
These methods equate salience with centrality, either to the input or some other
aggregate object (i.e. a cluster center or input centroid).
However, they rely
chiefly on
redundancy. When applied to an unfolding event, there may not exist enough
redundant content at the event onset for these methods to exploit.
Once the event onset has passed, however, the redundancy reduction of these 
methods is quite beneficial.

@@

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  Guo \textit{et al.} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster \cite{qi:temporal-summarization}.  The method uses a model of
sentence relevance and novelty in order to select appropriate updates.
Training data for regression targets is automatically generated from
retrospective Wikipedia data.  The system is evaluated on news documents
related to 197 natural and human disasters from 2009 to 2011 using variants of
Rouge modified to capture novelty, relevance, and comprehensiveness
\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
efficiency detect important updates during natural disasters
\cite{wang:update-summarization}.  The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than Guo \textit{et al.}.  The method is evaluated on
official press releases related to Hurricane Wilma  in 2005 using Rouge score
between the system summary and a manually generated target summary.

\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}

>>>>

First, centrality-focused approaches (including graph \cite{erkan2004lexrank},
cluster \cite{hatzivassiloglou2001simfinder}, and centroid \cite{radev2004centroid} methods) are very natural for retrospective analysis in
the sense that they let the data ``speak for itself.'' 
These methods equate salience with centrality, either to the input or some other
aggregate object (i.e. a cluster center or input centroid).
However, they rely
chiefly on
redundancy. When applied to an unfolding event, there may not exist enough
redundant content at the event onset for these methods to exploit.
Once the event onset has passed, however, the redundancy reduction of these 
methods is quite beneficial.

@@

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  \cite{qi:temporal-summarization} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster.  
%The method uses a model of
%sentence relevance and novelty in order to select appropriate updates.
%Training data for regression targets is automatically generated from
%retrospective Wikipedia data.  The system is evaluated on news documents
%related to 197 natural and human disasters from 2009 to 2011 using variants of
%ROUGE modified to capture novelty, relevance, and comprehensiveness
%\cite{lin2004rouge}.  
\cite{wang:update-summarization} present a clustering-based approach to
efficiency detect important updates during natural disasters.  
The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than \cite{qi:temporal-summarization}.  
Our system attempts to unify these system's approaches 
(predictive ranking and clustering respectively). 
%The method is evaluated on
%official press releases related to Hurricane Wilma  in 2005 using ROUGE score
%between the system summary and a manually generated target summary. 
%In addition
%to natural disasters, we evaluate our system on a variety of event types 
% such as terrorism and mass shootings (e.g., the 2012
%shooting in Aurora, Col.), accidents
%(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
%Arab spring). The size and scope of the events varies considerably. Hurricane
%Sandy, for example, affected multiple countries over weeks while the
%2012 Aurora shooting was contained to a single location spanning several hours.

%\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}

### 2015-02-28 results.tex
ck rev

Table~\ref{tab:rouge} shows our results for system output samples against the full summary of nuggets using ROUGE. 
%shows our results for the evaluation using ROUGE. 
%AP+Salience shows improvement on ngram precision and recall over the 
%baselines. 
This improvement is statistically significant for all ngram and longest common substring
precision, recall, and F-measure
at the $\alpha = .01$ level
using the Wilcoxon signed-rank test.

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure}

@@

\subsection{Expected Gain and Comprehensiveness}
\begin{figure}
  \includegraphics[]{nuggets-metrics.eps}
\caption{Expected Gain and Comprehensiveness performance.}
\label{fig:nperf}
\end{figure}

@@

Table~\ref{tab:farouge} shows the results of our feature ablation
tests. Removing the language models yields a statistically 
significant drop in both ngram recall and F-measure. 
Interestingly, removing the basic features leads to an
increase in both unigram and bigram precision; in the bigram
case this is enough to cause a statistically significant increase
in F-measure over the full model. In other words, the generic features
actually lead to an inferior model when we can incorporate more appropriate
domain specific features.
This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.

>>>>

Table~\ref{tab:rouge} shows our results for system output samples against the full summary of nuggets using ROUGE. 
%shows our results for the evaluation using ROUGE. 
%AP+Salience shows improvement on ngram precision and recall over the 
%baselines. 
This improvement is statistically significant for all ngram
precision, recall, and F-measures
at the $\alpha = .01$ level
using the Wilcoxon signed-rank test.

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\vspace{-14pt}
\end{figure}

@@

\subsection{Expected Gain and Comprehensiveness}
\begin{figure}
  \includegraphics[]{nuggets-metrics.eps}
\caption{Expected Gain and Comprehensiveness performance.}
\label{fig:nperf}
\vspace{-12pt}
\end{figure}

@@

Table~\ref{tab:farouge} shows the results of our feature ablation
tests. Removing the language models yields a statistically 
significant drop in both ngram recall and F-measure. 
Interestingly, removing the basic features leads to an
increase in both unigram and bigram precision; in the bigram
case this is enough to cause a statistically significant increase
in F-measure over the full model. In other words, the generic features
actually lead to an inferior model when we can incorporate more appropriate
domain specific features.
%This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.
